---
title: "topChef"

output: github_document

---

```{r, include = FALSE, message=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


## Introduction to topChef

topChef is a collection of data sets detailing events across all seasons of Top Chef US and Top Chef Masters US and one season of Top Chef Canada. It includes Chef information, challenge descriptions, challenge winners, episode information, guest judge names, and reward/prize information.

## Installation

Not yet on CRAN. So please use: *devtools::install.packages("celevitz/topChef")*
```{r InstallPackage , message=FALSE,warning=FALSE}
devtools::install_github("celevitz/topChef")
```

## References & Acknowlegements

Data were collected manually while watching each season of Top Chef. Additional data were collected from <https://en.wikipedia.org/wiki/Top_Chef>. My Top Chef data journey was inspired by <https://topchefstats.com/>.

Huge thanks to <https://github.com/doehm> for all his support!

```{r Library set up, message=FALSE, warning=FALSE,echo=FALSE,eval=TRUE}
library(topChef); library(tidyverse)
```

## Dataset overview

Across datasets, key joining variables include:

* `chef`
* `szn`
* `sznnumber`
* `series`
* `episode`

### Chef details

A tibble containing information about Chefs for each season they are in, including placement and gender. For some but not all seasons, there is also information on hometown, current city of residence, age, a flag for whether they are a person of color, and their occupation.

```{r Chef details table , eval=TRUE} 
chefdetails 
```

### Challenge descriptions
A tibble containing information about each challenge that the Chefs compete in.
```{r Challenge descriptions table , eval=TRUE} 
challengedescriptions 
```

### Challenge wins
A tibble containing win and loss data for each chef in each episode.
```{r Challenge results table , eval=TRUE} 
challengewins
```

### Judges
A tibble containing information about who were the guest judges for each challenge.
```{r Guest Judges table , eval=TRUE} 
judges 
```

### Rewards
A tibble containing information about rewards and prizes won by challenge.
```{r Rewards table , eval=TRUE} 
rewards
```

### Episode info
A tibble containing information about each episode.
```{r Episode information table , eval=TRUE} 
episodeinfo 
```

## Examples

### How many elimination challenge wins did Top Chef winners have?
#### Visualization
```{r EliminationWinsForTopChefsVisualization, eval=TRUE,echo=FALSE,message=FALSE} 
library(ggplot2)

chefdetails %>% 
  # Keep just winners and relevant variables (just the winning chef's name)
  filter(placement == 1) %>%
  select(chef, gender,series,szn) %>%
  # Bring on the challenge result data
  left_join(challengewins) %>%
  # Keep just elimination & sudden death quickfire challenges
  filter(challenge_type %in% c("Elimination","Quickfire Elimination","Sudden Death Quickfire")) %>%
  # Get the number of wins for each person
  group_by(series,szn,sznnumber,chef,gender) %>%
  filter(outcome %in% c("WIN","WINNER")) %>%
  summarise(wins=n()) %>%
  # Plot
  ggplot(aes(x=wins,y=chef,color=gender,fill=gender,shape=series)) +
    geom_point(size=4) +
    theme_minimal() + 
    # beautification
    scale_color_manual(values=c("#1170AA", "#A3ACB9")) +
    scale_shape_manual(values=c(0,17,1)) +
    scale_x_continuous(name="Number of elimination wins", 
                         breaks=seq(1,9,1),
                         limits=c(.8,9.5),
                         labels=seq(1,9,1)) +
    theme(panel.grid = element_blank(),
            title=element_text(size=12),
            axis.line.x=element_line(color="black"),
            axis.ticks.x=element_line(color="black"),
            axis.text.x=element_text(size=12),
            axis.title.x=element_text(size=12)) +
  labs(title="Number of elimination wins of Top Chefs"
       ,caption="I should have ordered by # of wins...")
    
```


#### Code
```{r EliminationWinsForTopChefsCode, eval=FALSE,echo=TRUE,message=FALSE} 
library(ggplot2)

chefdetails %>% 
  # Keep just winners and relevant variables (just the winning chef's name)
  filter(placement == 1) %>%
  select(chef, gender,series,szn) %>%
  # Bring on the challenge result data
  left_join(challengewins) %>%
  # Keep just elimination & sudden death quickfire challenges
  filter(challenge_type %in% c("Elimination","Quickfire Elimination","Sudden Death Quickfire")) %>%
  # Get the number of wins for each person
  group_by(series,szn,sznnumber,chef,gender) %>%
  filter(outcome %in% c("WIN","WINNER")) %>%
  summarise(wins=n()) %>%
  # Plot
  ggplot(aes(x=wins,y=chef,color=gender,fill=gender,shape=series)) +
    geom_point(size=4) +
    theme_minimal() + 
    # beautification
    scale_color_manual(values=c("#1170AA", "#A3ACB9")) +
    scale_shape_manual(values=c(0,17,1)) +
    scale_x_continuous(name="Number of elimination wins", 
                         breaks=seq(1,9,1),
                         limits=c(.8,9.5),
                         labels=seq(1,9,1)) +
    theme(panel.grid = element_blank(),
            title=element_text(size=12),
            axis.line.x=element_line(color="black"),
            axis.ticks.x=element_line(color="black"),
            axis.text.x=element_text(size=12),
            axis.title.x=element_text(size=12)) +
  labs(title="Number of elimination wins of Top Chefs"
       ,caption="I should have ordered by # of wins...")
    
```

### How do past seasons compare to how Season 20 is currently going?
#### Visualization
```{r Viz_IndexAllSeasons_SetUp, eval=TRUE,echo=FALSE,message=FALSE,include=FALSE} 
library(topChef); library(ggplot2); library(tidyverse)
 ## Get the index for all seasons
    
    allseasons <- weightedindex("US",1,10,7)
    for (season in seq(1,20,1)) {
      allseasons <- rbind(allseasons,weightedindex("US",season,10,7))
      
    }
```

```{r Viz_IndexAllSeasons, eval=TRUE,echo=FALSE,message=FALSE} 
    # drop unneeded variables
    allseasons <- allseasons[,c("chef","szn","sznnumber","placement","indexWeight")]
    
  ## Graph it
    # Add remaining Season 20 chefs to placement of 1.5
    # And, instead of showing that as placement, put it as current season
    allseasons$placement[is.na(allseasons$placement)] <- 1.5
    allseasons$label <- allseasons$placement
    allseasons$label[allseasons$placement == 1.5] <- "Current\nseason"
    
    # for sorting reasons, have the placement as a character
    allseasons$placementchar[allseasons$placement <= 9] <- 
      paste0("0",as.character(allseasons$placement[allseasons$placement <= 9]))
    allseasons$placementchar[allseasons$placement > 9] <- 
      as.character(allseasons$placement[allseasons$placement > 9])
    
  # Distribution of scores at this stage of the competition
    allseasons %>%
      ggplot(aes(x=szn,y=indexWeight) ) +
      geom_boxplot() +
      # add horizontal line at 0
      geom_rect(aes(xmin=0
                    ,xmax=20
                    ,ymin=-.5
                    ,ymax=0.5)
                ,fill="#ffbc69") +
      theme_minimal() +
      labs(title=paste0("Top Chef Weighted Index: 7 quickfires & 10 elimination challenges\ninto each season")
           ,subtitle="Comparing All Chefs Across All Seasons\n"
           ,caption="Scoring: Elimination win = 7 points. Elimination high = 3. Elimination low = -3. Eliminated = -7.\nQuickfire win = 4. Quickfire high = 2. Quickfire low = -2.\nData github.com/celevitz/topChef ||| Twitter @carlylevitz")+
      theme_minimal() +
      ylab("Index score") + xlab("") +
      theme(panel.grid = element_blank() 
            ,axis.text.x=element_text(angle=90)
            )
      
   # Top Four: Distribution of scores at this stage of the competition
    allseasons[allseasons$placement <= 4,]  %>%
      
      ggplot(aes(x=placement,y=indexWeight,group=label) ) +
      geom_boxplot() +
      # add horizontal line at 0
      geom_rect(aes(xmin=0
                    ,xmax=20
                    ,ymin=-.5
                    ,ymax=0.5)
                ,fill="#ffbc69") +
      theme_minimal() +
      labs(title=paste0("Top Chef Weighted Index: 7 quickfires & 10 elimination challenges\ninto each season")
           ,subtitle="Comparing Top Four Chefs Across All Seasons\n"
           ,caption="Scoring: Elimination win = 7 points. Elimination high = 3. Elimination low = -3. Eliminated = -7.\nQuickfire win = 4. Quickfire high = 2. Quickfire low = -2.\nData github.com/celevitz/topChef ||| Twitter @carlylevitz")+
      theme_minimal() +
      ylab("Index score") + xlab("") +
      theme(panel.grid = element_blank() 
            ,axis.text.x=element_text(angle=90)
            )
      
    

```
